---
title: "COVID19_Trend_analysis"
author: "David Ramirez"
date: 2/3/2021
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

The goal of this analysis is to answer the following questions using a dataset collected between 01/20/2020-06/01/20: 

1. Which countries have had the highest number of deaths due to COVID-19?
2. Which countries have had the highest number of positive cases against the number of tests?
3. Which countries have made the best effort in terms of the number of COVID-19 tests conducted related to their population?
4. Which countries were ultimately the most and least affected related to their population?


```{r Importing Packages and Libraries}
install.packages("tidyverse")
install.packages("tibble")
library(readr)
library(tibble)
library(dplyr)
library(ggplot2)
```

```{r Data Overview}
covid_df <- read_csv("covid19.csv")
dim(covid_df)
vector_cols <- colnames(covid_df)
glimpse(covid_df)
```

```{r Data Isolation}

covid_df_all_states <- filter(covid_df, Province_State == "All States")

covid_df_all_states_cummulative <- select(covid_df_all_states,
                                          Date, 
                                          Continent_Name, 
                                          Two_Letter_Country_Code, 
                                          positive, 
                                          hospitalized, 
                                          recovered, 
                                          death,total_tested
                                          )


covid_df_all_states_daily <- select(covid_df_all_states, 
                                    Date,
                                    Country_Region, 
                                    active,
                                    hospitalizedCurr,
                                    daily_tested,
                                    daily_positive
                                    )

glimpse(covid_df_all_states_cummulative)
glimpse(covid_df_all_states_daily)

```

```{r Countries w/ Greatest # of Deaths}
covid_df_all_states_cumulative_max <- covid_df_all_states_cummulative %>%
  group_by(Continent_Name,
           Two_Letter_Country_Code) %>%
  filter(death>0) %>%
  summarize(
    max_death = max(death)
  ) %>%
  arrange(-max_death) %>% 
  head(10)

show(covid_df_all_states_cumulative_max)


top_10_plot <- qplot(x=Two_Letter_Country_Code,
      y=max_death,
      colour=Continent_Name,
      data=covid_df_all_states_cumulative_max)
top_10_plot


death_top_3 <- c("United States(US)", "Italy(IT)", "United Kingdom(GB)")
```

```{r Highest Positive Cases vs. Tested Cases}

covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% 
  group_by(Country_Region) %>% 
  summarize(
    tested = sum(daily_tested), 
    postitive = sum(daily_positive), 
    active = sum(active), 
    hospitalized = sum(hospitalizedCurr)
  ) %>% 
  arrange(-tested)

show(covid_df_all_states_daily_sum)
covid_top_ten <- head(covid_df_all_states_daily_sum, 10)

countries <- covid_top_ten$Country_Region
tested_cases <- covid_top_ten$tested
positive_cases <- covid_top_ten$postitive
active_cases <- covid_top_ten$active
hospitalized_cases <- covid_top_ten$hospitalized

names(tested_cases) <- countries
names(positive_cases) <- countries
names( active_cases)<- countries
names(hospitalized_cases)<- countries

positive_tested_ratio = positive_cases/tested_cases
head(positive_tested_ratio,3)
positive_tested_top_3 <- c("United States"=0.1086, 
                           "Russia"=0.0385, 
                           "Italy" = 0.0615)

positive_tested_top_3
```



```{r Data Scaling to Pop. }
#population information retrieved from https://www.worldometers.info/world-population/population-by-country/

covid_mat <- cbind(tested_cases, positive_cases, active_cases, hospitalized_cases)
population <- c(331002651, 145934462, 60461826, 1380004385, 84339067, 37742154, 67886011, 25499884, 32971854, 37846611)

covid_mat <-covid_mat*100/population
show(covid_mat)


tested_cases_rank <- rank(covid_mat[,"tested_cases"])
positive_cases_rank <- rank(covid_mat[,"positive_cases"])
active_cases_rank <-  rank(covid_mat[,"active_cases"])
hospitalized_cases_rank <-  rank(covid_mat[,"hospitalized_cases"])

covid_mat_rank <- rbind(tested_cases_rank, 
                        positive_cases_rank, 
                        active_cases_rank, 
                        hospitalized_cases_rank)
covid_mat_rank
head(covid_mat_rank,1)

```



```{r Affected countries by rank}
colSums(covid_mat_rank[-1,])

best_effort_tested_cased_top_3 <- c("India", "United Kingdom", "Turkey")
most_affected_country <- c("Italy")
least_affected_country <-  c("India")
```


```{r Analysis Conclusion}
question_list <- list("Which countries have had the highest number of deaths due to COVID-19?", 
                      "Which countries have had the highest number of positive cases against the number of tests?", 
                      "Which countries have made the best effort in terms of the number of COVID-19 tests conducted related to their population?", 
                      "Which countries were ultimately the most and least affected related to their population?"
                      )

answer_list <-  list("Countries with the highest Death rate" = death_top_3,
                     "Countries with the highest Positive tested cases" = positive_tested_top_3,
                     "The best effort in testing related to the population" = best_effort_tested_cased_top_3,
                     "The most affected country related to its population" = most_affected_country,
                     "The least affected country related to its population" = least_affected_country
                     )

answer_list


```

## Including Plots

```{r 10 Highest Death Tolls, echo=FALSE}
top_10_plot

```


